The Design and Implementation of the Synchronous Language \CEU

### 2017-01-08 paper.tex
(>) timers

The \code{await} statement of \CEU supports wall-clock time and handles deltas 
automatically, resulting in more robust applications.
In the example in Figure~\ref{lst.timers}.a, suppose that after the first 
\code{await} request, the underlying system gets busy%
\footnote{Due to platform overheads (e.g., OS scheduling) and occasional slow
reactions, \code{await~10ms} means suspending a line of execution for
\emph{at least} 10 milliseconds.
}
and takes 15ms to notify 
\CEU.
The scheduler will notice that the \code{await~10ms} (ln. 2) has not only 
already expired, but is delayed with \code{delta=5ms}.
Then, the awaiting trail awakes, sets \code{v=1} (ln. 3), and invokes 
\code{await 1ms} (ln. 4).
As the current delta is still higher than the requested timeout (i.e. $5ms > 
1ms$), the trail is rescheduled for execution, now with \code{delta=4ms}.

@@

>>>>

The \code{await} statement of \CEU supports wall-clock time and handles deltas 
automatically. %, resulting in more robust applications.
In the example in Figure~\ref{lst.timers}.a, suppose that after the first 
\code{await} request, the underlying system gets busy%
\footnote{Due to platform overheads (e.g., OS scheduling) and occasional slow
reactions, \code{await~10ms} means suspending a line of execution for
\emph{at least} 10 milliseconds.
}
and takes 15ms to notify 
\CEU.
The scheduler will notice that the \code{await~10ms} (ln. 2) has not only 
already expired, but is delayed with \code{delta=5ms}.
Then, the awaiting trail awakes, sets \code{v=1} (ln. 3), and invokes 
\code{await 1ms} (ln. 4).
As the current delta is still higher than the requested timeout (i.e. $5ms > 
1ms$), the trail is rescheduled for execution, now with \code{delta=4ms}.

@@

The \code{await} statement for timers returns the current \emph{delta}, as
illustrated in Figure~\ref{lst.delta}.
Figure~\ref{lst.delta}.a is a variation of Figure~\ref{lst.timers}.a in which
the second \code{await} adds the current delta to annul it and guarantee that
\code{v} remains set to \code{1} during 1 millisecond.
Figure~\ref{lst.delta}.b deals with the possible case in which a small $1us$
timer in a loop never ``catches up'' with the external clock, resulting in a
\emph{delta} that increases indefinitely.

\begin{figure}
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}
var int v;
var int late = await 10ms;
v = 1;
await (late+1000)us;
v = 2;

.
\end{lstlisting}
\centering\small{\ax}
\end{minipage}
%
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}
loop do
    var int late = await 1us;
    if late > 1000 then     // I am late 1000x 1us?
        <...>               // abnormal behavior
    else
        <...>               // normal behavior
    end
end
\end{lstlisting}
\centering\small{\bx}
\end{minipage}
%\rule{8.5cm}{0.37pt}
\caption{ The \code{await} statement for timers returns the current
\emph{delta} in microseconds.
\label{lst.delta}
}
\end{figure}
